---
title: "p8105_hw6_jg4890"
author: "Jiayi Ge"
date: "2024-12-02"
output: github_document
---
# Problem 1

Load Libraries and Data
```{r}
library(tidyverse)
library(broom)
library(rsample)

weather_df <- rnoaa::meteo_pull_monitors(
  c("USW00094728"),
  var = c("PRCP", "TMIN", "TMAX"),
  date_min = "2017-01-01",
  date_max = "2017-12-31"
) %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10
  ) %>%
  select(name, id, everything())
```

Bootstrap Analysis
```{r}
set.seed(123)

bootstrap_samples <- bootstraps(weather_df, times = 5000)

bootstrap_metrics <- function(split) {

  sample_data <- analysis(split)
  
  model <- lm(tmax ~ tmin, data = sample_data)
  
  r_squared <- glance(model)$r.squared
  
  coefs <- tidy(model) %>% pull(estimate)
  
  log_beta_product <- log(coefs[1] * coefs[2])
  
  return(tibble(r_squared = r_squared, log_beta_product = log_beta_product))
}

bootstrap_results <- bootstrap_samples %>%
  mutate(metrics = map(splits, bootstrap_metrics)) %>%
  unnest(metrics)
```

Plot Distributions

```{r}
ggplot(bootstrap_results, aes(x = r_squared)) +
  geom_histogram(binwidth = 0.01, fill = "blue", alpha = 0.7) +
  labs(
    title = "Bootstrap Distribution of R-Squared",
    x = "R-Squared",
    y = "Frequency"
  )

ggplot(bootstrap_results, aes(x = log_beta_product)) +
  geom_histogram(binwidth = 0.02, fill = "green", alpha = 0.7) +
  labs(
    title = "Bootstrap Distribution of log(beta_0 * beta_1)",
    x = "log(beta_0 * beta_1)",
    y = "Frequency"
  )
```

Confidence Intervals
```{r}
conf_intervals <- bootstrap_results %>%
  summarize(
    r_squared_lower = quantile(r_squared, 0.025),
    r_squared_upper = quantile(r_squared, 0.975),
    log_beta_product_lower = quantile(log_beta_product, 0.025),
    log_beta_product_upper = quantile(log_beta_product, 0.975)
  )
print(conf_intervals)

```

# Problem 2
```{r}
library(tidyverse)
library(broom)
library(purrr)
library(dplyr)

url <- "https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv"
homicides <- read_csv(url)

homicides <- homicides %>%
  mutate(victim_age = as.numeric(victim_age)) %>%
  filter(!(city == "Dallas" & state == "TX"),
         !(city == "Phoenix" & state == "AZ"),
         !(city == "Kansas City" & state == "MO"),
         !(city == "Tulsa" & state == "AL")) %>%
  filter(victim_race %in% c("White", "Black")) %>%
  filter(!is.na(victim_age)) %>%
  mutate(city_state = paste(city, state, sep = ", "),
         solved_binary = ifelse(disposition == "Closed by arrest", 1, 0))

print(homicides)
```

Logistic regression for Baltimore, MD
```{r}
baltimore <- homicides %>%
  filter(city_state == "Baltimore, MD")

baltimore_model <- glm(
  solved_binary ~ victim_age + victim_sex + victim_race,
  family = "binomial",
  data = baltimore
)
baltimore_results <- broom::tidy(baltimore_model, conf.int = TRUE) %>%
  filter(term == "victim_sexMale") %>%
  mutate(
    adj_odds_ratio = exp(estimate),
    conf.low = exp(conf.low),
    conf.high = exp(conf.high)
  )

print(baltimore_results)
```

Logistic regression for all cities
```{r}
city_results <- homicides %>%
  group_by(city_state) %>%
  nest() %>%
  mutate(
    model = map(data, ~ glm(
      solved_binary ~ victim_age + victim_sex + victim_race,
      family = "binomial",
      data = .
    )),
    results = map(model, ~ broom::tidy(., conf.int = TRUE) %>%
      filter(term == "victim_sexMale") %>%
      mutate(
        adj_odds_ratio = exp(estimate),
        conf.low = exp(conf.low),
        conf.high = exp(conf.high)
      ))
  ) %>%
  unnest(results)

city_results %>%
  ggplot(aes(x = reorder(city_state, adj_odds_ratio), y = adj_odds_ratio)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  coord_flip() +
  labs(
    title = "Adjusted Odds Ratios for Solving Homicides",
    x = "City",
    y = "Adjusted Odds Ratio (Male vs Female)"
  ) +
  theme_minimal()
```
The plot visualizes the adjusted odds ratios (ORs) for solving homicides involving male versus female victims across various U.S. cities. The horizontal line segments indicate the 95% confidence interval for the OR. 

#### Trend in Odds Ratios:

**Majority Below 1**: Most cities exhibit ORs below 1, suggesting that homicides involving male victims are generally less likely to be solved compared to those involving female victims.

**Variation Across Cities**: However, there is considerable variation in the magnitude of these differences across cities. Some cities, like Baltimore, show a pronounced difference, while others, like New York, have a more modest difference.

#### Confidence Intervals:

**Width and Precision**: The width of the confidence intervals varies across cities. Narrower intervals, like those for Baltimore, indicate more precise estimates. Wider intervals, such as those for Albuquerque and Stockton, suggest greater uncertainty in the OR estimate.

**Statistical Significance**: If the confidence interval for a city does not include 1, it indicates a statistically significant difference in the odds of solving homicides between male and female victims.

#### Potential Factors Influencing ORs:

**Local Law Enforcement Practices**: Differences in investigative techniques, resource allocation, and prioritization of cases across police departments may influence the likelihood of solving homicides.
Socioeconomic Factors: Factors like poverty, crime rates, and community engagement can impact the resources available for investigations and the cooperation of witnesses.

**Victim Characteristics**: Beyond sex, factors like age, race, and socioeconomic status may also influence the likelihood of a homicide being solved.

# Problem3 
```{r}
library(tidyverse)
library(modelr)
library(broom)
library(crossval)
library(purrr)

birthweight <- read_csv("./data/birthweight.csv")
birthweight <- birthweight %>%
  mutate(
    babysex = factor(babysex, labels = c("Male", "Female")),
    malform = factor(malform, labels = c("Absent", "Present")),
    frace = factor(frace),
    mrace = factor(mrace)
  ) %>%
  drop_na()

model_proposed <- lm(bwt ~ babysex + bhead + blength + delwt + gaweeks + wtgain, data = birthweight)

birthweight <- birthweight %>%
  add_predictions(model_proposed, var = "predicted") %>%
  add_residuals(model_proposed, var = "residuals")

residual_plot <- birthweight %>%
  ggplot(aes(x = predicted, y = residuals)) +
  geom_point() +
  geom_smooth() +
  geom_hline(yintercept = 0, color = "red", linetype = "dashed") +
  labs(title = "Residuals vs Fitted Values", x = "Fitted Values", y = "Residuals") +
  theme_minimal()

print(residual_plot)
```

#### Modeling process

The dataset was cleaned by converting categorical variables (`babysex`, `frace`, `mrace`, `malform`) into factors to allow appropriate statistical modeling. Missing data were dropped to ensure analysis integrity. The dependent variable, `bwt` (birth weight), remained numeric for regression analysis. 

A multiple linear regression model was developed to predict birth weight using hypothesized influential factors: `bhead`, `blength`, `gaweeks`, `delwt`, and `wtgain`. These variables were chosen based on biological plausibility and domain knowledge about factors influencing fetal growth.

Residuals and predicted values were computed using `add_predictions()` and `add_residuals()`. The residual plot revealed residuals scattered around zero, indicating no severe violations of linearity or heteroscedasticity. In addition, slight deviations were noted at extreme fitted values, may hint at potential model refinements.

**Comparison models**
```{r}
# Model 1: Length at birth and gestational age (main effects only)
comparison_model_1 <- lm(bwt ~ blength + gaweeks, data = birthweight)

# Model 2: Head circumference, length, sex, and all interactions (including three-way)
comparison_model_2 <- lm(bwt ~ bhead * blength * babysex, data = birthweight)
```

```{r}
cv_df <- crossv_mc(birthweight, 100) |>
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  )

cv_res <- cv_df |>
  mutate(
    model_proposed = map(train, ~lm(bwt ~ babysex + bhead + blength + delwt + gaweeks + wtgain, data = .x)),
    comparison_model_1 = map(train, ~lm(bwt ~ blength + gaweeks, data = .x)),
    comparison_model_2 = map(train, ~lm(bwt ~ bhead * blength * babysex, data = .x)),
    rmse_proposed = map2_dbl(model_proposed, test, ~rmse(.x, .y)),
    rmse_model1 = map2_dbl(comparison_model_1, test, ~rmse(.x, .y)),
    rmse_model2 = map2_dbl(comparison_model_2, test, ~rmse(.x, .y))
  )

cv_res |>
  select(starts_with("rmse")) |>
  pivot_longer(
    everything(),
    names_to = "model",
    values_to = "rmse",
    names_prefix = "rmse_"
  ) |>
  ggplot(aes(x = model, y = rmse)) +
  geom_violin() +
  labs(
    title = "rmse comparison of models",
    x = "model",
    y = "rmse"
  )
```

The violin plot visualizes the distribution of RMSE values for three different regression models across 100 cross-validation folds. The proposed model demonstrates the lowest RMSE values, indicating superior predictive performance compared to Model 1 (using only length at birth and gestational age) and Model 2 (using head circumference, length, sex, and all their interactions). This may suggest that the inclusion of additional relevant factors in the proposed model enhances its predictive accuracy for birthweight.
